---
title: "AAI1001_poster_draft"
format: 
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    section-depth: 2
    embed-resource: true
editor: visual
author: 
  - Toh Shao Ern
  - Kuick Si Qi
  - Wong Khin Foong
  - Sebestian Fernandez
  - Yu Haotong
---

# AAI1001 Poster Draft

## Load Library

```{r include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE)
if (!require("sf")) install.packages("sf", dependencies = TRUE)
if (!require("rnaturalearth")) install.packages("rnaturalearth", dependencies = TRUE)
if (!require("countrycode")) install.packages("countrycode", dependencies = TRUE)
if (!require("ggrepel")) install.packages("ggrepel", dependencies = TRUE)
if (!require("readxl")) install.packages("readxl", dependencies = TRUE)
if (!require("knitr")) install.packages("knitr", dependencies = TRUE)
if (!require("dplyr")) install.packages("dplyr", dependencies = TRUE)
if (!require("viridis")) install.packages("viridis", dependencies = TRUE)
if (!require("gridExtra")) install.packages("gridExtra", dependencies = TRUE)
if (!require("grid")) install.packages("grid", dependencies = TRUE)
if (!require("ggplot2")) install.packages("ggplot2", dependencies = TRUE)
```

```{r include=FALSE}

library(readxl)
library(dplyr)
library(knitr)
library(tidyverse) # load dplyr, ggplot2, stringr, etc.
library(sf) # working with geographic simple features in R
library(rnaturalearth) # World map data from Natural Earth
library(countrycode) # get ISO code from country names
library(ggrepel) # "ggplot2" extension for overlapping text labels
library(viridis) # color palettes for "ggplot2")
library(gridExtra) # arrange multiple grid-based plots on a page
library(grid) # grid graphics system for R
library(ggplot2) # create elegant data visualizations using the Grammar of Graphics
```

## Load Data

```{r}
df <- read_excel("data/DataForTable2.1.xls")
df
```

## Data Understanding

```{r}
summary(df)
head(df)
```

## List and Identify Missing Countries

```{r}
#| code-fold: true
#| code-summary: "Show the total list of Countries in the world"
countries <- c(
  "Afghanistan", "Albania", "Algeria", "Andorra", "Angola", "Antigua and Barbuda", 
  "Argentina", "Armenia", "Australia", "Austria", "Azerbaijan", "Bahamas", "Bahrain", 
  "Bangladesh", "Barbados", "Belarus", "Belgium", "Belize", "Benin", "Bhutan", "Bolivia", 
  "Bosnia and Herzegovina", "Botswana", "Brazil", "Brunei", "Bulgaria", "Burkina Faso", 
  "Burundi", "Cabo Verde", "Cambodia", "Cameroon", "Canada", "Central African Republic",
  "Chad", "Chile", "China", "Colombia", "Comoros", "Congo", "Costa Rica", "Croatia", 
  "Cuba", "Cyprus", "Czechia", "Denmark", "Djibouti", "Dominica", "Dominican Republic", 
  "East Timor", "Ecuador", "Egypt", "El Salvador", "Equatorial Guinea", "Eritrea", 
  "Estonia", "Eswatini", "Ethiopia", "Fiji", "Finland", "France", "Gabon", "Gambia", 
  "Georgia", "Germany", "Ghana", "Greece", "Grenada", "Guatemala", "Guinea", "Guinea-Bissau", 
  "Guyana", "Haiti", "Honduras", "Hungary", "Iceland", "India", "Indonesia", "Iran", 
  "Iraq", "Ireland", "Israel", "Italy", "Jamaica", "Japan", "Jordan", "Kazakhstan", 
  "Kenya", "Kiribati", "Kuwait", "Kyrgyzstan", "Laos", "Latvia", "Lebanon", "Lesotho", 
  "Liberia", "Libya", "Liechtenstein", "Lithuania", "Luxembourg", "Madagascar", "Malawi", 
  "Malaysia", "Maldives", "Mali", "Malta", "Marshall Islands", "Mauritania", "Mauritius", 
  "Mexico", "Micronesia", "Moldova", "Monaco", "Mongolia", "Montenegro", "Morocco", 
  "Mozambique", "Myanmar", "Namibia", "Nauru", "Nepal", "Netherlands", "New Zealand", 
  "Nicaragua", "Niger", "Nigeria", "North Korea", "North Macedonia", "Norway", "Oman", 
  "Pakistan", "Palau", "State of Palestine", "Panama", "Papua New Guinea", "Paraguay", "Peru", 
  "Philippines", "Poland", "Portugal", "Qatar", "Romania", "Russia", "Rwanda", "Saint Kitts and Nevis", 
  "Saint Lucia", "Saint Vincent and the Grenadines", "Samoa", "San Marino", "Sao Tome and Principe", 
  "Saudi Arabia", "Senegal", "Serbia", "Seychelles", "Sierra Leone", "Singapore", "Slovakia", 
  "Slovenia", "Solomon Islands", "Somaliland region", "South Africa", "South Korea", "South Sudan", 
  "Spain", "Sri Lanka", "Sudan", "Suriname", "Sweden", "Switzerland", "Syria", "Taiwan Province of China", 
  "Tajikistan", "Tanzania", "Thailand", "Togo", "Tonga", "Trinidad and Tobago", "Tunisia", 
  "Türkiye", "Turkmenistan", "Tuvalu", "Uganda", "Ukraine", "United Arab Emirates", "United Kingdom", 
  "United States", "Uruguay", "Uzbekistan", "Vanuatu", "Vatican City", "Venezuela", "Vietnam", 
  "Yemen", "Zambia", "Zimbabwe"
)
length(countries)
```

```{r}
# how many countries are missing in the dataset use Country name
missing_countries <- setdiff(countries, unique(df$"Country name"))
missing_countries

```

We can see that there are 34 countries missing from the dataset.

## Data Cleaning and Transformation

As we will be only using data from 2021 - 2023, we will filter out the data from other years. We will also check for missing data and fill in the missing values with the mean of the available data from each country.

### Filter data for years 2021 - 2023

```{r}
df <- df %>%
  filter(year %in% c(2021, 2022, 2023))
head(df)
```

### Renaming headers and keeping only the necessary columns

```{r}
# Rename the columns
colnames(df) <- c("country", "year", "score", colnames(df)[4:ncol(df)])

# Keep only the specified columns
df_filtered <- df[, c("country", "year", "score")]
head(df_filtered)
```

### Include Missing Countries with NA Scores

```{r}
# List of all countries in your data
all_countries <- unique(df_filtered$country)

# List of years to ensure each country has entries for each year
years <- c(2021, 2022, 2023)

# Create a data frame with all combinations of countries and years
complete_df <- expand.grid(country = all_countries, year = years)

# Add country codes to the complete data frame
complete_df$country_code <- countrycode(complete_df$country, origin = "country.name", destination = "iso3c")

# Merge the complete data frame with the original data
df_combined <- left_join(complete_df, df_filtered, by = c("country", "year"))

# Calculate the mean score for each country
country_means <- df_combined %>%
  group_by(country) %>%
  summarize(mean_score = mean(score, na.rm = TRUE))

# Merge the country means back into the combined data
df_combined <- left_join(df_combined, country_means, by = "country")

# Fill any remaining NA values with the mean score of the respective country
df_combined <- df_combined %>%
  mutate(score = ifelse(is.na(score), mean_score, score))

# Check for any remaining missing values
missing_scores <- df_combined %>% filter(is.na(score))
if (nrow(missing_scores) > 0) {
  warning("There are still countries with missing scores.")
}

# Inspect the data for missing scores
print(missing_scores)
```

```{r}
region <- read_csv("data/world-regions-according-to-the-world-bank.csv")
head(region)

df_combined_region <- left_join(df_combined, region, by = c("country_code" = "Code"))
head(df_combined_region)
# find empty rows in "World Region according to the World Bank"


missing_region <- df_combined_region %>% filter(is.na(`World Region according to the World Bank`))
print(missing_region)

df_combined_region$`World Region according to the World Bank`[which(df_combined_region$country == "Taiwan Province of China")] <- "East Asia and Pacific"

missing_region <- df_combined_region %>% filter(is.na(`World Region according to the World Bank`))
print(missing_region)

df_combined_region$`World Region according to the World Bank`[which(df_combined_region$`World Region according to the World Bank` == "North America")] <- "Latin America and Caribbean"
df_combined_region$`World Region according to the World Bank`[which(df_combined_region$`World Region according to the World Bank` == "Latin America and Caribbean")] <- "America and Caribbean"

```

#Create violin plot

```{r}
ggplot(df_combined_region, aes(y = `World Region according to the World Bank`, x = mean_score, fill = `World Region according to the World Bank`)) +
  geom_violin(trim = FALSE) +  # Create horizontal violin plots
  geom_boxplot(width = 0.1, fill = "white", color = "black", orientation = "y") +  # Overlay boxplots on the violins horizontally
  scale_fill_viridis(discrete = TRUE, guide = FALSE) +  # Use viridis color scale for discrete data and remove legend
  labs(title = "Happiness Scores by World Region (2021-2023)", y = "World Region", x = "Happiness Score") +
  theme_minimal() +  # Use a minimal theme
  theme(
    plot.title = element_text(hjust = 0.5)  # Center the plot title
  )

```

## Data Visualization

### World Map

```{r}
# Get world data
world <- ne_countries(scale = "small", returnclass = "sf")

# Filter data for each year
df_2021 <- df_combined %>% filter(year == 2021)
df_2022 <- df_combined %>% filter(year == 2022)
df_2023 <- df_combined %>% filter(year == 2023)

# Convert country names to country codes
df_2021$country_code <- countrycode(df_2021$country, origin = "country.name", destination = "iso3c")
df_2022$country_code <- countrycode(df_2022$country, origin = "country.name", destination = "iso3c")
df_2023$country_code <- countrycode(df_2023$country, origin = "country.name", destination = "iso3c")

# Merge world data with each year's data
world_data_2021 <- left_join(world, df_2021, by = c("iso_a3" = "country_code"))
world_data_2022 <- left_join(world, df_2022, by = c("iso_a3" = "country_code"))
world_data_2023 <- left_join(world, df_2023, by = c("iso_a3" = "country_code"))

# Check for missing data after the join
missing_world_2021 <- world_data_2021 %>% filter(is.na(score))
missing_world_2022 <- world_data_2022 %>% filter(is.na(score))
missing_world_2023 <- world_data_2023 %>% filter(is.na(score))

print(missing_world_2021)
print(missing_world_2022)
print(missing_world_2023)

# Plot data for 2021
ggplot() +
  geom_sf(data = world_data_2021, aes(fill = score)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "World Map of Happiness - 2021", fill = "Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_sf(crs = "+proj=robin")

# Plot data for 2022
ggplot() +
  geom_sf(data = world_data_2022, aes(fill = score)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "World Map of Happiness - 2022", fill = "Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_sf(crs = "+proj=robin")

# Plot data for 2023
ggplot() +
  geom_sf(data = world_data_2023, aes(fill = score)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "World Map of Happiness - 2023", fill = "Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_sf(crs = "+proj=robin")
```

```{r}
# Create bins for the scores with 0.5 intervals
bins <- seq(0, 10, by = 0.5)
labels <- paste0(bins[-length(bins)], "-", bins[-1])
df_combined <- df_combined %>%
  mutate(score_bin = cut(score, breaks = bins, labels = labels, include.lowest = TRUE, right = FALSE))


# Check for any remaining missing values
missing_scores <- df_combined %>% filter(is.na(score))
if (nrow(missing_scores) > 0) {
  warning("There are still countries with missing scores.")
}

# Inspect the data for missing scores
print(missing_scores)

# Get world data
world <- ne_countries(scale = "small", returnclass = "sf")

# Filter data for each year
df_2021 <- df_combined %>% filter(year == 2021)
df_2022 <- df_combined %>% filter(year == 2022)
df_2023 <- df_combined %>% filter(year == 2023)

# Merge world data with each year's data
world_data_2021 <- left_join(world, df_2021, by = c("iso_a3" = "country_code"))
world_data_2022 <- left_join(world, df_2022, by = c("iso_a3" = "country_code"))
world_data_2023 <- left_join(world, df_2023, by = c("iso_a3" = "country_code"))

# Check for missing data after the join
missing_world_2021 <- world_data_2021 %>% filter(is.na(score))
missing_world_2022 <- world_data_2022 %>% filter(is.na(score))
missing_world_2023 <- world_data_2023 %>% filter(is.na(score))

print(missing_world_2021)
print(missing_world_2022)
print(missing_world_2023)

# Plot data for 2021
ggplot() +
  geom_sf(data = world_data_2021, aes(fill = score), color = "black") +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Score") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "World Map of Happiness - 2021", fill = "Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_sf(crs = "+proj=robin")

# Plot data for 2022
ggplot() +
  geom_sf(data = world_data_2022, aes(fill = score), color = "black") +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Score") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "World Map of Happiness - 2022", fill = "Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_sf(crs = "+proj=robin")

# Plot data for 2023
ggplot() +
  geom_sf(data = world_data_2023, aes(fill = score), color = "black") +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Score") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "World Map of Happiness - 2023", fill = "Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_sf(crs = "+proj=robin")
```

## Top 5, Middle 5, Bottom 5 Countries Bar Graph

```{r}
library(ggplot2)
library(dplyr)
library(viridis)
library(scales)

# Calculate the mean score for each country
country_means <- df_combined %>%
  group_by(country) %>%
  summarize(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(desc(mean_score))

# Assign colors based on scores
assign_colors <- function(scores, palette = "plasma", direction = 1) {
  normalized_scores <- rescale(scores, to = c(0, 1))
  colors <- viridis_pal(option = palette, direction = direction)(length(unique(normalized_scores)))
  score_colors <- setNames(colors[rank(unique(normalized_scores), ties.method = "first")], unique(normalized_scores))
  return(score_colors[as.character(normalized_scores)])
}

# Apply colors globally
country_means$color <- assign_colors(country_means$mean_score)

# Calculate the top, middle, and bottom countries
top_5 <- country_means %>% top_n(5, mean_score) %>% mutate(category = "Top 5")
middle_start <- ceiling((nrow(country_means) - 5) / 2)
middle_end <- middle_start + 4
middle_5 <- country_means %>% slice(middle_start:middle_end) %>% mutate(category = "Middle 5")
bottom_5 <- country_means %>% top_n(-5, mean_score) %>% mutate(category = "Bottom 5")

# Combine the data and set the category order
combined_data <- bind_rows(top_5, middle_5, bottom_5) %>%
  mutate(category = factor(category, levels = c("Top 5", "Middle 5", "Bottom 5")))

# Create the faceted plot
plot_combined <- ggplot(combined_data, aes(x = reorder(country, mean_score), y = mean_score, fill = color)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(mean_score, 2), hjust = -0.1)) +
  coord_flip() +
  scale_fill_identity() +
  labs(x = "Country", y = "Mean Score") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0, 10)) +
  facet_wrap(~category, scales = "free_y", ncol = 1)

# Print the combined plot
print(plot_combined)


```

## Distribution Chart of Scores

```{r}

data = df_combined


# Fit a normal distribution to the data
mu <- mean(data$score, na.rm = TRUE)
std <- sd(data$score, na.rm = TRUE)
median <- median(data$score, na.rm = TRUE)
mode <- as.numeric(names(sort(table(data$score), decreasing = TRUE)[1]))  # Mode calculation

# Create histogram data
hist_data <- data %>%
  filter(!is.na(score)) %>%
  ggplot(aes(x = score)) +
  geom_histogram(bins = 20, color = "white", alpha = 0.75, aes(y = ..count.., fill = ..x..)) +
  scale_fill_viridis(option = "plasma", name = "Score") +
  labs(title = "Distribution of Happiness Scores (2021-2023)",
       x = "Happiness Score",
       y = "Number of Countries") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold", color = "#333333", margin = margin(b = 20)),
        axis.title.x = element_text(size = 14, color = "#333333"),
        axis.title.y = element_text(size = 14, color = "#333333"),
        axis.text = element_text(size = 12, color = "#333333"),
        panel.grid.major.y = element_line(linetype = "dotted", size = 0.5, color = "grey"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

# Filter out zero count bins
hist_data <- ggplot_build(hist_data)$data[[1]] %>%
  filter(count > 0)

# Create the histogram with the filtered data
histogram <- ggplot(hist_data, aes(x = x, y = count, fill = x)) +
  geom_bar(stat = "identity", color = "white", alpha = 0.75) +
  scale_fill_viridis(option = "plasma", name = "Score") +
  labs(title = "Distribution of Happiness Scores (2021-2023)",
       x = "Happiness Score",
       y = "Number of Countries") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold", color = "#333333", margin = margin(b = 20)),
        axis.title.x = element_text(size = 14, color = "#333333"),
        axis.title.y = element_text(size = 14, color = "#333333"),
        axis.text = element_text(size = 12, color = "#333333"),
        panel.grid.major.y = element_line(linetype = "dotted", size = 0.5, color = "grey"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

# Add count labels on top of each bar
histogram <- histogram +
  geom_text(aes(label = count), vjust = -0.5, size = 4)

# Define the range for x and compute y for the normal distribution
x <- seq(min(data$score, na.rm = TRUE), max(data$score, na.rm = TRUE), length.out = 100)
y <- dnorm(x, mean = mu, sd = std)

# Increase the height of the curve by scaling
scale_factor <- 150  # Adjust this value as needed
y <- y * scale_factor

# Add the normal distribution curve
histogram <- histogram +
  geom_line(data = data.frame(x = x, y = y), aes(x = x, y = y), color = "#006400", size = 1, linetype = "dotted") +
  geom_vline(xintercept = mu, color = "red", linetype = "dashed", size = 1)

# Find the y-value of the mean line on the normal distribution curve
mean_y <- dnorm(mu, mean = mu, sd = std) * scale_factor

# Extract plot limits
xlims <- ggplot_build(histogram)$layout$panel_params[[1]]$x.range
ylims <- ggplot_build(histogram)$layout$panel_params[[1]]$y.range

# Create a table with statistics
stats_table <- textGrob(
  label = paste("Median: ", sprintf("%.2f", median), "\n",
                "Mode: ", sprintf("%.2f", mode), "\n",
                "Standard Deviation: ", sprintf("%.2f", std)),
  gp = gpar(fontsize = 12, col = c("blue", "green", "purple")),
  hjust = 0,
  vjust = 1
)

# Create the mean annotation separately
mean_annotation <- textGrob(
  label = sprintf("Mean: %.2f", mu),
  gp = gpar(fontsize = 12, col = "red"),
  hjust = 1,
  vjust = 1
)

# Add the statistics table to the top-left corner and the mean annotation to the top-left of the mean line
histogram <- histogram +
  annotation_custom(grob = stats_table, xmin = xlims[1], xmax = xlims[1] + (xlims[2] - xlims[1]) * 0.2, ymin = ylims[2], ymax = ylims[2] - (ylims[2] - ylims[1]) * 0.2) +
  annotation_custom(grob = mean_annotation, xmin = mu - (xlims[2] - xlims[1]) * 0.05, xmax = mu, ymin = ylims[2] - (ylims[2] - ylims[1]) * 0.1, ymax = ylims[2])

# Print the histogram
print(histogram)


```
